TSELM: Target speaker extraction using discrete tokens and language models

### 2024-09-06 src/paper.tex
update

\begin{table*}
    \caption{The performance of different systems on Libri2Mix testset. For TSELM-L-Hybrid  
    and TSELM, we compare the speaker similarity with the discretized target speech 
    (Target-discrete-WavLM) instead of the target speech due to the observation that 
    the process of discretization already losses the speaker information (0.653 for 
    Target-discrete-WavLM). We use "\_d" to denote it.  }
    \setlength{\tabcolsep}{12pt} % Adjust column spacing
    \renewcommand{\arraystretch}{1.2}
    \begin{center}
        \begin{tabular}{ccccccccc}
            \Xhline{2\arrayrulewidth} % Bold top line
            \multirow{2}{*}{System} & \multirow{2}{*}{Category} & \multirow{2}{*}{Model type} & \multirow{2}{*}{Model size} & \multicolumn{3}{c}{ DNSMOS $\uparrow$} & \multirow{2}{*}{dWER $\downarrow$} & \multirow{2}{*}{Spk Sim $\uparrow$} \\
            \cline{5-7}
              &                   &                             &                             & SIG     & BAK     & OVL    &                       &                          \\ 
            \hline
            Mixture                 & - & -                           & -                           & 3.39    & 3.14    & 2.68   & 79.3\%                & -                        \\
            Target-Discrete   & G & -                          & 330M                           & 3.57    & 4.10    & 3.32   & 11.3\%                & 0.653                    \\
            \hline
            Spex+                   &  D & Continuous                  & 11M                         & 3.36    & 3.76    & 2.98   & 19.3\%                & 0.923                    \\
            \hline
            Continuous-WavLM-L6   & G    & Continuous                  & 220M                        & 3.56    & 4.06    & 3.27   & 14.7\%                & 0.877                    \\
            TSELM-L-Hybrid    & G       & Hybrid                      & 468M                        & 3.50    & 4.06    & 3.22   & 19.8\%                & 0.924\_d                 \\
            TSELM-S-NoConcat           & G       & Discrete                    & 354M                        & 3.47    & 4.03    & 3.19   & 69.6\%                & 0.868\_d                 \\
            \hline
            TSELM-S           & G       & Discrete                    & 354M                        & 3.50    & 4.07    & 3.24   & 28.1\%                & 0.892\_d                 \\
            TSELM-M          & G        & Discrete                    & 389M                        & 3.49    & 4.05    & 3.22   & 28.4\%                & 0.901\_d                 \\
            TSELM-L         & G         & Discrete                    & 468M                        & 3.49    & 4.05    & 3.22   & 27.1\%                & 0.905\_d \\
            \Xhline{2\arrayrulewidth} % Bold top line              
            \end{tabular}
            \linebreak
            \label{main_exp}
      \end{center}
    %   \vspace{-10pt}
  \end{table*}

@@

\subsection{Baseline models}
Baseline models are reflected in Table \ref{main_exp}. 
TSELM is compared with Spex+, a discriminative separation model \cite{spex_plus} trained on Libri2Mix \cite{librimix}. Since the discretization might lose some 
speaker information compared with continuous methods. We compare the speaker 
similarity of TSELM with the target audio produced by discretizing the clean audio, which is denoted by Target-Discrete, and it is the upper bound of our model  
performance. Besides TSELM, we did three main experiments 
denoted by Continuous-WavLM-L6, TSELM-L-Hybrid and TSELM-S-NoConcat using the same training data. 
For Continuous-WavLM-6, we directly
passed the embeddings of the 6th hidden layer output of WavLM to the cross attention  
and LM. Concatenation strategy is stilled applied to the mixed speech. Mean Square Error (MSE) loss is applied between the output embeddings and the clean 
embeddings. HiFi-GAN in \cite{knn_vc} is used to reconstruct the audio. For 
TSELM-L-Hybrid, inspired by MaskSR \cite{mask_sr}, we kept discretizing the reference 
speech while utilizing the continuous embeddings from the reference speech. We call 
it hybrid instead of fully discretized because the mixture speech has continuous 
features instead of discrete features. For TSELM-S-NoConcat, we utilize the TSELM-S model architecture but without concatenation strategy.

>>>>

\begin{table*}
    \caption{The performance of different systems on Libri2Mix testset. For TSELM-L-Hybrid  
    and TSELM, we compare the speaker similarity with the discretized target speech 
    (Target-discrete-WavLM) instead of the target speech due to the observation that 
    the process of discretization already losses the speaker information (0.653 for 
    Target-discrete-WavLM). We use "\_d" to denote it.  }
    \setlength{\tabcolsep}{12pt} % Adjust column spacing
    \renewcommand{\arraystretch}{1.2}
    \begin{center}
        \begin{tabular}{ccccccccc}
            \Xhline{2\arrayrulewidth} % Bold top line
            \multirow{2}{*}{System} & \multirow{2}{*}{Category} & \multirow{2}{*}{Model type} & \multirow{2}{*}{Model size} & \multicolumn{3}{c}{ DNSMOS $\uparrow$} & \multirow{2}{*}{dWER $\downarrow$} & \multirow{2}{*}{Spk Sim $\uparrow$} \\
            \cline{5-7}
              &                   &                             &                             & SIG     & BAK     & OVL    &                       &                          \\ 
            \hline
            Mixture                 & - & -                           & -                           & 3.39    & 3.14    & 2.68   & 79.3\%                & -                        \\
            Target-Discrete   & G & -                          & 330M                           & 3.57    & 4.10    & 3.32   & 11.3\%                & 0.653                    \\
            \hline
            Spex+                   &  D & Continuous                  & 11M                         & 3.36    & 3.76    & 2.98   & 19.3\%                & 0.923                    \\
            \hline
            \hline
            Continuous-WavLM-L6   & G    & Continuous                  & 220M                        & 3.56    & 4.06    & 3.27   & 14.7\%                & 0.877                    \\
            TSELM-L-Hybrid    & G       & Hybrid                      & 468M                        & 3.50    & 4.06    & 3.22   & 19.8\%                & 0.924\_d                 \\
            TSELM-S-NoConcat           & G       & Discrete                    & 354M                        & 3.47    & 4.03    & 3.19   & 69.6\%                & 0.868\_d                 \\
            \hline
            TSELM-S           & G       & Discrete                    & 354M                        & 3.50    & 4.07    & 3.24   & 28.1\%                & 0.892\_d                 \\
            TSELM-M          & G        & Discrete                    & 389M                        & 3.49    & 4.05    & 3.22   & 28.4\%                & 0.901\_d                 \\
            TSELM-L         & G         & Discrete                    & 468M                        & 3.49    & 4.05    & 3.22   & 27.1\%                & 0.905\_d \\
            \Xhline{2\arrayrulewidth} % Bold top line              
            \end{tabular}
            \linebreak
            \label{main_exp}
      \end{center}
    %   \vspace{-10pt}
  \end{table*}

@@

\subsection{Baseline models}
Baseline models are presented in Table \ref{main_exp}. 
TSELM is compared with Spex+ \cite{spex_plus}, a discriminative separation model trained on Libri2Mix \cite{librimix}. Since discretization can potentially lose speaker information compared to continuous methods. We compute the speaker 
similarity of output from TSELM with the target audio produced by discretizing the clean audio, denoted by Target-Discrete, representing the upper bound of our model  
performance. Besides TSELM, we conduct three main experiments 
denoted by Continuous-WavLM-L6, TSELM-L-Hybrid and TSELM-S-NoConcat using the same training data. 
For Continuous-WavLM-6, we directly
passed the embeddings of the 6th hidden layer output of WavLM to the cross attention  
and LM. Concatenation strategy is stilled applied to the mixed speech. Mean Square Error (MSE) loss is applied between the output embeddings and the clean 
embeddings. HiFi-GAN in \cite{knn_vc} is used to reconstruct the audio. For 
TSELM-L-Hybrid, inspired by MaskSR \cite{mask_sr}, we kept discretizing the reference 
speech while utilizing the continuous embeddings from the reference speech. We call 
it hybrid instead of fully discretized because the mixture speech has continuous 
features instead of discrete features. For TSELM-S-NoConcat, we utilize the TSELM-S model architecture but without concatenation strategy.
